Formal Modelling of Predator Preferences using Molecular Gut-Content Analysis

### 2015-09-24 paper/main.tex
line 367 corrected

Beyond the statistical assumptions, sampling of prey availability should be given careful attention.  To survey prey of interest, we relied on pitfall trapping, which is a reasonable method to quantify potential prey encounters of sit-and-pursue predators, such as wolf spiders \citep{Uetz:1999}.  In general, however, human-made traps are liable to collect prey at rates unequal to that of the predator of interest.  Confounded with researchers' abilities to make sufficiently efficient traps are a host of environmental variables.  Researchers should consider these potential sampling issues and attempt to determine, a priori, how well their traps will characterize prey availability according to the biology of their focal predator.  Mathematically, these issues will directly affect the parameter $c_{st} = \lambda_{st}/\gamma_{st}$.  When human-made traps are relatively more efficient at catching prey than are the predators, $c_{st}$ will be relatively smaller.  Conversely, when the traps are relatively less efficient, $c_{st}$ will be relatively larger.  Unfortunately, these factors can be difficult to control, statistically, practically, or otherwise, and are to some degree unavoidable.  Mark-recapture methods might be implemented to adjust for trap efficiency and estimate the true density of each prey species.  However, recapture rates are likely to be low and this type of data is beyond what is regularly collected in these types of experiments.

@@

\nocite{Wickham:2011a,Wickham:2011,Bischl:2014,Roualdes:2014}
\bibliographystyle{spbasic}
\bibliography{refs}
\end{document}

>>>>

Beyond the statistical assumptions, sampling of prey availability should be given careful attention.  To survey prey of interest, we relied on pitfall trapping, which is a reasonable method to quantify potential prey encounters of sit-and-pursue predators, such as wolf spiders \citep{Uetz:1999}.  In general, however, human-made traps are liable to collect prey at rates unequal to that of the predator of interest.  Confounded with researchers' abilities to make sufficiently efficient traps are a host of environmental variables.  Researchers should consider these potential sampling issues and attempt to determine, a priori, how well their traps will characterize prey availability according to the biology of their focal predator.  If such issues are steady over time, then mathematically the parameter $c_{st} = \lambda_{st}/\gamma_{st}$ is directly affected.  When human-made traps are relatively and consistently more efficient at catching prey than are the predators, $c_{st}$ will be relatively smaller.  Conversely, when the traps are relatively and consistenly less efficient, $c_{st}$ will be relatively larger.  Unfortunately, these factors can be difficult to control, practically, or otherwise, and are to some degree unavoidable.  Mark-recapture methods might be implemented to adjust for trap efficiency and estimate the true density of each prey species.  However, recapture rates are likely to be low and this type of data is beyond what is regularly collected in these types of experiments.

If issues of trap efficiency or abundance are suspected to change across sampling times, we offer an alternative parameterization, details of which are found in Appendix~\ref{sec:efficiency}.  The ratio $c_{st} = \lambda_{st} / \gamma_{st}$ implicitly treats $\gamma_{st}$ as an index proportional to true abundance.  This assumption is valid if the traps maintain efficiency across all occassions, but otherwise is incorrect.  Rewriting $c_{st}$ as $k_{st}d_t$, such that $\sum_{s=1}^S k_{st} = 1$ for all $t$, allows for trap efficiency to vary across time.  The parameter $d_t$ captures changes in the absolute relationship between $\lambda_{st}$ and $\gamma_{st}$ which may be due to actual changes in abundance or to changes in trap efficiency.  $k_{st}$ then appropriately measures the ratio of consumption to encounter rates, taking into account changes in abundance or trap efficiency.

@@

\appendix
\section{Varying Trap Efficiency}
\label{sec:efficiency}

When there is concern about trap efficiency varying across time,  we recommend the following reparameterization of $c_{st} = \lambda_{st}/\gamma_{st}$.  Let

\[ k_{st}d_t = \frac{\lambda_{st}}{\gamma_{st}}, \]

\noindent such that $\sum_{s=1}^S k_{st} = 1$, for all $t$.  The parameter $d_t$ allows for changes in trap efficiency or in actual abundance over time.  If such changes are believed to be present, this reparameterization allows for a more natural interpretation.  The parameter $k_{st}$ models relative consumption to encounter rates, as expected.

Since this reparameterization amounts to a linear transformation of the parameters, the asymptotic distribution for $\hat{d}_t$ can be evaluated directly.  Because of the constraint $\sum_{s=1}^S k_{st} = 1$, we find that $d_t = \sum_{s=1}^S c_{st}$ or written in matrix notation  $\mathbf{d} = \mathbf{1}' \mathbf{c}$, where $\mathbf{d} \in \mathbb{R}^T$.  The asymptotic distribution of $\mathbf{\hat{d}}$ follows immediately, $\mathbf{\hat{d}} \stackrel{\cdot}{\sim} \mathcal{N}(\mathbf{1}'\mathbf{d}, \mathbf{1}'\Sigma\mathbf{1})$, where $\Sigma$ is the covariance matrix of the asymptotic distribution of $\mathbf{\hat{c}}$.

More importantly, we can find the asymptotic distribution of $\mathbf{\hat{k}}_t = (k_{1t}, \ldots, k_{St})' \in \mathbb{R}^S$.  We make use of the delta method \citep{Serfling:2001}.  The asymptotic variance relies on the derivatives, $d k_{rt} / d c_{st}$.  Let $c_{\cdot t} = \sum_{s=1}^S c_{st}$.  Because $k_{st} = c_{st}/ c_{\cdot t}$, we can find the necessary derivatives in two cases

\[
  \frac{d k_{rt}}{d c_{st}} =
\begin{cases}
  \frac{-c_{rt}}{c_{\cdot t}^2}, & r \ne s \\
  \frac{c_{\cdot t} - c_{st}}{c_{\cdot t}^2}, & r = s.
\end{cases}
\]

\noindent  Thus, the asymptotic distribution of $\mathbf{\hat{k}}_{t}$ is $\mathcal{N}(\frac{c_{st}}{c_{\cdot t}}, (\frac{d \mathbf{k}_t}{d \mathbf{c}_t}) \Sigma (\frac{d \mathbf{k}_t}{d \mathbf{c}_t})')$. 
\nocite{Wickham:2011a,Wickham:2011,Bischl:2014,Roualdes:2014}
\bibliographystyle{spbasic}
\bibliography{refs}
\end{document}
